--- 
title: "Data Visualization with R"
author: "Mohammed Amimul Ehasan Shakib"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This repository is a comprehensive collection of useful charts and visualizations created using R, a language renowned for its capabilities in statistical computing, graphical data analysis, and scientific research."
---

# Charts

This repository offers a comprehensive collection of useful charts and visualizations created using R.

### Example 1: Scatter Plot


```{r }
library(ggplot2)

# ggplot2

ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  ggtitle("Scatter Plot (ggplot2)") +
  xlab("Weight") +
  ylab("Miles per Gallon")
```

### Example 2: Line Plot

```{r }
#Line Plot without package
plot(1:10, 1:10, type = "l", main = "Line Plot (Base R)", xlab = "X", ylab = "Y")

#Line Plot with package ggplot2
ggplot(data.frame(x = 1:10, y = 1:10), aes(x = x, y = y)) +
  geom_line() +
  ggtitle("Line Plot (ggplot2)") +
  xlab("X") +
  ylab("Y")
```


### Example 3: Bar Plot

```{r }
#Bar Plot without package
data <- table(iris$Species)
barplot(data, main = "Bar Plot (Base R)", xlab = "Species", ylab = "Count")

#Bar Plot with package ggplot2
 
ggplot(data.frame(Species = levels(iris$Species)), aes(x = Species)) +
  geom_bar() +
  ggtitle("Species Counts") +
  xlab("Species") +
  ylab("Count")
```



### Example 4: Histogram

```{r }
#Histogram without package
hist(mtcars$mpg, main = "Histogram (Base R)", xlab = "Miles per Gallon", ylab = "Frequency" )

#Histogram with package ggplot2
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 2, fill = "blue", color = "black") +
  ggtitle("Histogram (ggplot2)") +
  xlab("Miles per Gallon") +
  ylab("Frequency")
```



### Example 5: Box Plot

```{r }
#Box Plot without package
boxplot(mtcars$mpg, main = "Box Plot (Base R)")

#Box Plot with package ggplot2
ggplot(mtcars, aes(y = mpg)) +
  geom_boxplot() +
  ggtitle("Box Plot (ggplot2)") +
  ylab("Miles per Gallon")
```



### Example 6: Violin Plot

```{r }
#Violin Plot with package ggplot2
ggplot(iris, aes(x = Species, y = Sepal.Length, fill = Species)) +
  geom_violin() +
  ggtitle("Violin Plot (ggplot2)") +
  xlab("Species") +
  ylab("Sepal Length")
```



### Example 7: Heatmap (pheatmap)

```{r }
# Install and load the pheatmap package (if not already installed)
#install.packages("pheatmap")

library(pheatmap)

# Compute the correlation matrix for mtcars
correlation_matrix <- cor(mtcars)

# Create a heatmap
pheatmap(correlation_matrix,
         color = colorRampPalette(c("blue", "white", "red"))(100),
         main = "Heatmap of mtcars",
         cluster_rows = FALSE, cluster_cols = FALSE,
         fontsize = 10) # Adjust font size
```



### Example 8: Time Series Plot (Base R)

```{r }
data("AirPassengers")
ts_plot <- ts(AirPassengers, start = c(1949, 1), frequency = 12)
plot(ts_plot, main = "Time Series Plot (Base R)", xlab = "Year", ylab = "Passengers")
```



### Example 9: 3D Scatter Plot (scatterplot3d)

```{r }
library(scatterplot3d)
scatterplot3d(mtcars$mpg, mtcars$hp, mtcars$wt, main = "3D Scatter Plot (scatterplot3d)",
              xlab = "MPG", ylab = "Horsepower", zlab = "Weight", color = "blue")
```



### Example 10: Pie Chart (Base R)

```{r }
#Pie Chart without package
data <- table(iris$Species)
pie(data, main = "Pie Chart (Base R)")
 
```



### Example 11: Radar Chart (fmsb)

```{r }
library(fmsb) # Load the necessary library

data <- scale(iris[, 1:4]) # Select the first four columns of the iris dataset and scale them
data <- data.frame(data, Species = iris$Species) # Create a data frame with row names as species

radarchart(data, axistype = 1, title = "Radar Chart (fmsb)") # Create a radar chart
```



### Example 12: Density Plot (ggplot2)

```{r }
ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_density(alpha = 0.5) +
  ggtitle("Density Plot (ggplot2)") +
  xlab("Sepal Length") +
  ylab("Density")
```



### Example 13: Bar Chart with Error Bars (ggplot2)

```{r }
set.seed(123)
x <- 1:10
y <- x + rnorm(length(x))
error <- runif(length(x), 0.1, 0.3)
df <- data.frame(x, y, error)
ggplot(df, aes(x = x, y = y)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = y - error, ymax = y + error), width = 0.2) +
  ggtitle("Bar Chart with Error Bars (ggplot2)") +
  xlab("X") +
  ylab("Y")
```



### Example 14: Parallel Coordinate Plot (ggplot2)

```{r }
ggplot(iris, aes(x = Species, y = Sepal.Length, color = Species)) +
  geom_path() +  # Use geom_path for parallel coordinates
  ggtitle("Parallel Coordinates Plot (ggplot2)") +
  xlab("Species") +
  ylab("Sepal Length")
```



### Example 15: Sunburst Chart (plotly)

```{r }
library(plotly)
data <- data.frame(
  id = c("A", "B", "C", "D", "E", "F"),
  parent = c("", "", "", "A", "A", "D"),
  value = c(10, 14, 12, 6, 8, 4)
)
fig <- plot_ly(data, ids = ~id, labels = ~id, parents = ~parent, values = ~value,
               type = 'sunburst')
fig <- fig %>% layout(title = "Sunburst Chart (plotly)")
fig
```



### Example 16: Mosaic Plot (mosaic)

```{r }
# Load the mosaic package
library(mosaic)
library(vcd)

# Load the Arthritis dataset
data("Arthritis")

# Create a mosaic plot
mosaic(~ Improved + Treatment, data = Arthritis,
       main = "Mosaic Plot (Arthritis)")
```



### Example 17: Word Cloud (wordcloud2)

```{r }
library(wordcloud2)
# Create a sample word frequency data frame
wordcloud_data <- data.frame(
  word = c("crime1", "crime2", "crime3", "crime4", "crime5"),
  freq = c(20, 15, 10, 5, 3)
)

# Create a word cloud
wordcloud2(wordcloud_data, size = 0.5,
           color = "random-dark", backgroundColor = "black")


```



### Example 18: Hexbin Plot (hexbinplot)

```{r }
library(hexbin)
data("iris")
hexbinplot(Petal.Length ~ Petal.Width | Species, data = iris,
           xlab = "Petal Width", ylab = "Petal Length",
           main = "Hexbin Plot (hexbinplot)")

```



### Example 19: Ternary Plot (vcd)

```{r }
library(vcd)

# Create a sample matrix with 3 columns
# Replace this with your own data
ternary_data <- matrix(runif(300), ncol = 3)
# Create a ternary plot
ternaryplot(ternary_data,
            main = "Ternary Plot (vcd)")
```



### Example 20: Candlestick Chart (quantmod)

```{r }
# Install and load the required package
library(quantmod)

# Get stock price data for AAPL (Apple Inc.) from Yahoo Finance
getSymbols("AAPL", src = "yahoo", from = "2022-01-01", to = "2022-12-31")

# Create a candlestick chart
chartSeries(AAPL, type = "candlesticks", theme = "white",
            name = "Apple Inc. (AAPL) Stock Prices",
            subset = "2022-01-01::2022-12-31")

```



### Example 21: Wind Rose Plot (openair)

```{r }
#Install the required dependencies
library(maps)
library(openair)

# Set a seed for reproducibility
set.seed(123)

# Generate random wind direction (0 to 360 degrees) and wind speed data
directions <- runif(100, 0, 360)
speeds <- runif(100, 0, 10)

# Create a data frame with wind data
wind_data <- data.frame(direction = directions, speed = speeds)

# Create a wind rose plot
windRose(mydata = wind_data,
         ws = "speed", wd = "direction",
         key.header = "Wind Rose Plot (openair)")

```



### Example 22: Sunflower Plot (graphics)

```{r }
set.seed(123)
x <- rnorm(100)
y <- rnorm(100)
sunflowerplot(x, y, main = "Sunflower Plot (graphics)")

```



### Example 23: Cluster Dendrogram (dendextend)

```{r }
library(dendextend)
data("mtcars")
dend <- as.dendrogram(hclust(dist(mtcars)))
dend <- color_branches(dend, k = 3)
plot(dend, main = "Cluster Dendrogram (dendextend)")
```



### Example 24: Network Plot (igraph)

```{r }
library(igraph)
set.seed(123)
nodes <- data.frame(name = letters[1:10])
edges <- data.frame(
  from = sample(nodes$name, 20, replace = TRUE),
  to = sample(nodes$name, 20, replace = TRUE)
)
g <- graph_from_data_frame(d = edges, vertices = nodes)
plot(g, layout = layout_with_kk(g), vertex.label = nodes$name, main = "Network Plot (igraph)")
```



### Example 25: Radial Plot (circlize)

```{r }
# Load the circlize package
library(circlize)

# Load the mtcars dataset
data("mtcars")

# Compute the correlation matrix
cor_matrix <- cor(mtcars)

# Set row names to match column names
rownames(cor_matrix) <- colnames(cor_matrix)

# Create a chord diagram
chordDiagram(cor_matrix, transparency = 0.5, annotationTrack = "grid",
             preAllocateTracks = 1)

# Add a title to the plot
title(main = "Radial Plot (circlize)")
#----------- before this line self learning
```



### Example 26: Mosaic Plot (vcd)

```{r }
library(vcd)
data("Arthritis")
mosaic(~ Improved + Treatment, data = Arthritis, shade = TRUE,
       main = "Mosaic Plot (vcd)")
```



### Example 27: Star Plot (fmsb)

```{r }
# Install and load the 'fmsb' package
#install.packages("fmsb")
library(fmsb)

# Load the 'mtcars' dataset
data("mtcars")

# Create a binary variable 'mpg01' based on median
mtcars$mpg01 <- ifelse(mtcars$mpg > median(mtcars$mpg), "high", "low")

# Create a star plot (radar chart)
radarchart(mtcars[, 1:7], title = "Star Plot (fmsb)")
```



### Example 28: Beeswarm Plot (ggbeeswarm)

```{r }
# Install and load the 'ggbeeswarm' package
#install.packages("ggbeeswarm")
library(ggbeeswarm)

# Load the 'mtcars' dataset
data("mtcars")

# Create a beeswarm plot
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_beeswarm() +
  ggtitle("Beeswarm Plot (ggbeeswarm)") +
  xlab("Cylinders") +
  ylab("Miles per Gallon")
```



### Example 29: QQ Plot (base R)

```{r }
qqnorm(rnorm(100), main = "QQ Plot (base R)")
qqline(rnorm(100))
```



### Example 30: Parallel Coordinates Plot (ggplot2)

```{r }
library(ggplot2)
data("iris")
ggplot(iris, aes(x = Species, y = Sepal.Length, color = Species)) +
  geom_path() +
  ggtitle("Parallel Coordinates Plot (ggplot2)") +
  xlab("Species") +
  ylab("Sepal Length")
```



### Example 31: Venn Diagram (VennDiagram)

```{r }
library(VennDiagram)
venn.plot <- venn.diagram(
  x = list(A = 1:100, B = 51:150, C = 101:200),
  category.names = c("A", "B", "C"),
  filename = NULL
)
grid.draw(venn.plot)
```



### Example 32: Surface Plot (plotly)

```{r }
library(plotly)
x <- seq(-10, 10, length = 100)
y <- seq(-10, 10, length = 100)
z <- outer(x, y, function(x, y) sin(sqrt(x^2 + y^2)) / sqrt(x^2 + y^2))
plot_ly(z = ~z, type = "surface") %>% layout(title = "Surface Plot (plotly)")
```



### Example 33: Contour Plot (base R)

```{r }
x <- seq(-2 * pi, 2 * pi, length = 100)
y <- seq(-2 * pi, 2 * pi, length = 100)
f <- function(x, y) cos(x) * sin(y)
z <- outer(x, y, f)
contour(x, y, z, main = "Contour Plot (base R)", xlab = "X", ylab = "Y")

```


